Automatic Detection of Bad Programming Habits in Scratch

### 2014-10-18 slides/slides.tex
Some changes here and there

\section{About}

@@

\begin{center}
\Huge Are bad programming habits a common practice in the Scratch community?
\end{center}

@@

\begin{itemize}
  \item Educators teaching to code
  \item Students learning to code
  \item Programming learning tools developers
\end{itemize}

@@

  \begin{columns}[T]
    \begin{column}{1\textwidth}
     \begin{block}{Learning to code with Scratch}
\begin{itemize}
  \item Success to teach both basic and advanced programming concepts
  \item Bad programming habits detected
  \item Lack of automatic tools to check for \textit{correctness}
  \item Hairball: Lint-inspired Static Analysis of Scratch Projects
\end{itemize}
    \end{block}
    \end{column}

@@

%--------------------------------------------------------
\begin{frame}
\frametitle{Bad programming habits with Scratch: objects naming}

@@

\begin{center}
Sprites naming
\end{center}
\end{frame}

%--------------------------------------------------------
\begin{frame}
\frametitle{Bad programming habits with Scratch: repetition code}

\begin{figure}[t!]
\begin{center}
\includegraphics[width=6cm,height=2.7cm]{figs/CodeRepetition1.png}
\end{center}
\label{fig:repetition1}
\end{figure}

\begin{center}
Repetition code
\end{center}

\begin{figure}[t!]
\begin{center}
\includegraphics[width=6cm,height=2.7cm]{figs/CodeRepetition2.png}
\end{center}
\label{fig:repetition2}
\end{figure}

\begin{center}
Creating blocks to avoid repetition code
\end{center}

@@

We developed two plug-ins for Hairball to automatically detect bad programming habits:
\begin{itemize}
  \item convention.SpriteNaming
  \item duplicate.DuplicateScripts
\end{itemize}

@@

\begin{table}
\begin{center}
  \begin{tabular}{ | l | r | r | r |}
   \hline
              & Default names & Duplicated scripts & Defined blocks \\ \hline\hline
    Projects & 79 & 62 & 17 \\ \hline
    Mean & 5.94 & 7.23 & 1.11 \\ \hline
    Median & 3 & 2 & 0 \\ \hline
    Maximum & 67 & 71 & 25 \\
    \hline    
  \end{tabular}
\end{center}
\caption{Analysis of 100 ramdonly downloaded projects}
\label{table:results}
\end{table}
\end{frame}

@@

\begin{itemize}
  \item Extend the scope of the study developing new plug-ins
  \item Analyze dataset with 5 years data from Scratch website
  \item Dr.Scratch: http://drscratch.programamos.es
\end{itemize}
\vspace{\baselineskip}
\vspace{\baselineskip}
\hfill{\Tiny Background picture: Simon Cunningham }

>>>>

\section{FIE 2014 - eMadrid Session}

@@

\begin{center}
\Huge {\bf Are bad programming habits a common practice in the Scratch community?}
\end{center}

@@

\begin{itemize}
  \item Educators teaching how to code
  \item Students learning to program
  \item Developers of programming learning tools
\end{itemize}

@@

  \begin{columns}[T]
    \begin{column}{1\textwidth}
     \begin{block}{Learning to code with Scratch}
\begin{itemize}
  \item Scratch has shown to be successfull in teaching basic and advanced programming concepts
  \item However, bad programming habits have been detected
  \item There are no automatic tools to check for \textit{correctness}
  \item \texttt{Hairball}: \texttt{lint}-inspired static analysis of Scratch projects
\end{itemize}
    \end{block}
    \end{column}

@@

%--------------------------------------------------------
\begin{frame}
\frametitle{Bad programming habits with Scratch (I)}

@@

\begin{center}
\emph{Bad}/default naming of sprites
\end{center}
\end{frame}

%--------------------------------------------------------
\begin{frame}
\frametitle{Bad programming habits in Scratch (and II)}

  \begin{columns}[T]
    \begin{column}{0.5\textwidth}

     \begin{block}{Example of repeated code}
\begin{figure}[t!]
\begin{center}
\includegraphics[width=5.4cm,height=2.5cm]{figs/CodeRepetition1.png}
\end{center}
\label{fig:repetition1}
\end{figure}
     \end{block}
    \end{column}
    \begin{column}{0.5\textwidth}
     \begin{block}{Solution to avoid repeated code}
\begin{figure}[t!]
\begin{center}
\includegraphics[width=5.4cm,height=2.5cm]{figs/CodeRepetition2.png}
\end{center}
Blocks should be created to avoid repetition of code
\label{fig:repetition2}
\end{figure}
     \end{block}

    \end{column}
  \end{columns}

@@

  \begin{columns}[T]
    \begin{column}{0.8\textwidth}
     \begin{block}{\vspace{0.1cm}\center{We have developed two plug-ins for Hairball to automatically detect bad programming habits}\vspace{0.1cm}}
\vspace{0.3cm}
\begin{enumerate}
  \item convention.SpriteNaming
\vspace{0.15cm}
  \item duplicate.DuplicateScripts
\end{enumerate}
\vspace{0.4cm}
    \end{block}
    \end{column}
  \end{columns}

@@

\begin{table}
\begin{center}
  \begin{tabular}{ | l | c | c | c |}
   \hline
              & Default names & Duplicated scripts & Defined blocks \\ \hline\hline
    Projects & 79 & 62 & 17 \\ \hline
    Mean & 5.94 & 7.23 & 1.11 \\ \hline
 22   Median & 3 & 2 & 0 \\ \hline
    Maximum & 67 & 71 & 25 \\
    \hline    
  \end{tabular}
\end{center}
\caption{Analysis of 100 ramdonly downloaded Scratch projects}
\label{table:results}
\end{table}
\end{frame}

@@

\begin{enumerate}
  \item Extend the scope of the study developing new plug-ins
  \item Analyze dataset with 5 years of data from the Scratch website
  \item Dr. Scratch (alpha version): http://drscratch.programamos.es
\end{enumerate}
\vspace{\baselineskip}
\vspace{\baselineskip}
\hfill{\Tiny Background picture: Simon Cunningham }
